Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-03-18 strang/paper.tex
Make sure all variables are defined

Simulations of stellar flows require solving the equations of
hydrodynamics coupled with a nuclear reaction network.  The equations
of hydrodynamics with reacting sources are:
\begin{align}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho X_k \Ub) &= \rho \omegadot_k \\
\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
\end{align}
where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions,
$\omegadot_k$ are the species creation rates (from reactions), $p$ is the pressure, $E$ is the
total energy / unit mass, and $\Sdot$ is the nuclear energy generation rate.

When we are reacting, we can look at internal energy:
\begin{equation}
\rho \frac{De}{Dt} + p \nabla \cdot \Ub = \rho \Sdot
\end{equation}
or alternately, we can evolve the temperature, which we can write as:
\begin{equation}
\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
\end{equation}
(this form neglects the change in temperature due to composition
changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
equilibrium).

@@

\begin{figure}[t]
\centering
\plotone{test}
\caption{\label{thefigure} Convergence of fluid quantities as a function of resolution
for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$ with $c_v$ held fixed, and evolving $(X_k, e)$.}
\end{figure}

>>>>

Simulations of stellar flows require solving the equations of
hydrodynamics coupled with a nuclear reaction network.  The equations
of hydrodynamics with reacting sources are:
\begin{align}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho X_k \Ub) &= \rho \omegadot_k \\
\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
\end{align}
where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions,
$\omegadot_k$ are the species creation rates (from reactions), $p$ is the pressure, $E$ is the
specific total energy, and $\Sdot$ is the nuclear energy generation rate.

When we are reacting, we can look at internal energy
\begin{equation}
\rho \frac{De}{Dt} + p \nabla \cdot \Ub = \rho \Sdot,
\end{equation}
where $e$ is the specific internal energy or alternately, we can evolve the temperature, 
which we can write as
\begin{equation}
\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot,
\end{equation}
where $c_v$ is the specific heat capacity at constant volume, $T$ is the temperature and $e_\rho$ is the derivative of 
the specific internal energy with respect to density (this form neglects the change in 
temperature due to composition
changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
equilibrium).

@@

\begin{figure}[t]
\centering
\plotone{test}
\caption{\label{thefigure} Convergence of fluid quantities as a function of resolution
for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$ with $c_v$ held fixed, and evolving $(X_k, e)$. The dotted lines show ideal first and second order convergence.}
\end{figure}

### 2021-03-18 xrb2/paper.tex
Make sure all variables are defined

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. We saw in \citet{flame_wave1} that 
increasing the neutron star rotation rate reduces the horizontal lengthscale of the flame. An 
estimate of this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius 
may be thought of as the scale over which the balance between the Coriolis force and horizontal 
pressure gradient becomes important, and is approximately given by
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. \added{If we take $g \sim 3.6 \times 10^{12}~\mathrm{cm}~\mathrm{s}^{-2}$ and $H_0\sim 10^3~\mathrm{cm}$, then for the $500~\mathrm{Hz}$ system we find the Rossby length to approximately be $L_R \sim 1.2 \times 10^5~\mathrm{cm}$. This estimate is likely to be an overestimate of the true value of the Rossby length in our simulations, as we saw that the flame is confined on a much smaller length scale.} In Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, we use $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates. Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be 
reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). \added{As the natural aspect ratio makes it hard to see the flame structure, we show the energy generation rate for the $\tt{F1000}$ simulation with the vertical extent stretched in Figure~\ref{fig:flame_stretch}.}  From Equation 
\ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
$1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater confinement from 
the Coriolis force should reduce the horizontal extent of the flame by a similar factor. However, 
the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our 
simulations we see that this scaling does not work so well for all rotation rates. The simulations 
seem to follow the theoretical scaling described in Equation \ref{eqn:rossby} more closely at 
higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.

>>>>

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. We saw in \citet{flame_wave1} that 
increasing the neutron star rotation rate reduces the horizontal lengthscale of the flame. An 
estimate of this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius 
may be thought of as the scale over which the balance between the Coriolis force and horizontal 
pressure gradient becomes important, and is approximately given by
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. \added{If we take $g \sim 1.5 \times 10^{14}~\mathrm{cm}~\mathrm{s}^{-2}$ and $H_0\sim 10^3~\mathrm{cm}$, then for the $500~\mathrm{Hz}$ system we find the Rossby length to approximately be $L_R \sim 1.2 \times 10^5~\mathrm{cm}$. This estimate is likely to be an overestimate of the true value of the Rossby length in our simulations, as we saw that the flame is confined on a much smaller length scale.} In Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, we use $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates. Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be 
reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). \added{As the natural aspect ratio makes it hard to see the flame structure, we show the energy generation rate for the $\tt{F1000}$ simulation with the vertical extent stretched in Figure~\ref{fig:flame_stretch}.}  From Equation 
\ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
$1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater confinement from 
the Coriolis force should reduce the horizontal extent of the flame by a similar factor. However, 
the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our 
simulations we see that this scaling does not work so well for all rotation rates. The simulations 
seem to follow the theoretical scaling described in Equation \ref{eqn:rossby} more closely at 
higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
